---
output:
  word_document: default
  html_document: default
---
##Speed Dating Experient

#Import data
```{r}
library(readr)
Sdate <- read_csv("~/Academic/MS Statistics/STAT 6509 Regression Modeling/Project_Speed_Dating/Speed Dating Data.csv")
```

#Examine the data
```{r}
str(Sdate)
```
#Select the dataset we are interested in
```{r}
which(names(Sdate)%in%c("dec", "attr","sinc","intel","fun","amb","shar","age","samerace","gender","imprelig","goal","date","go_out","exphappy"))
Sdate11<-Sdate[,c(3,15,34,42,46:48,68,98:104)]
```

#Check the missing values
```{r}
library(Amelia)
missmap(Sdate11,main = "Missing values vs observed")
```

# Explore study field
```{r}
fields.cd <- c(
  "Law",
  "Math",
  "Social Science, Psychologist" ,
  "Medical Science, Pharmaceuticals, and Bio Tech",
  "Engineering",
  "English/Creative Writing/ Journalism",
  "History/Religion/Philosophy",
  "Business/Econ/Finance",
  "Education, Academia",
  "Biological Sciences/Chemistry/Physics",
  "Social Work" ,
  "Undergrad/undecided" ,
  "Political Science/International Affairs" ,
  "Film",
  "Fine Arts/Arts Administration",
  "Languages",
  "Architecture",
  "Other"
)


# Find number of men/women on each study field
library(dplyr)
fields <- Sdate[!is.na(Sdate$field_cd),] %>%
  group_by(gender, field_cd) %>%
  summarise(
    my.n = n()
  )

# Plot study fields repartition
library(ggplot2)
ggplot(fields, aes(x = field_cd, y = my.n, fill = factor(gender))) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_discrete(name = "Gender") +
  xlab("Field") + ylab("Count") + ggtitle("Study fields repartition") +
  scale_x_continuous(labels = fields.cd, breaks = 1:18) +
  coord_flip()



```

#Explore the intended career
```{r}
career.cd <- c(
  "Lawyer",
  "Academic/Research", 
  "Psychologist", 
  "Doctor/Medicine",
  "Engineer", 
  "Creative Arts/Entertainment",
  "BankingBusiness/CEO/Admin",
  "Real Estate",
  "International/Humanitarian Affairs",
  "Undecided" ,
  "Social Work",
  "Speech Pathology",
  "Politics",
  "Pro sports/Athletics",
  "Other",
  "Journalism",
  "Architecture"
)

careers <- Sdate[!is.na(Sdate$career_c),] %>%
  group_by(gender, career_c) %>%
  summarise(
    my.n = n()
  )

ggplot(careers, aes(x = career_c, y = my.n, fill = factor(gender))) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_discrete(name = "Gender") +
  xlab("Career") + ylab("Count") + ggtitle("Careers repartition") +
  scale_x_continuous(labels = career.cd, breaks = 1:17) +
  coord_flip()

```

#Explore the age range
```{r}
Sdate$age[is.na(Sdate$age)] <- mean(Sdate$age,na.rm=T)
Sdate$gender<-as.factor(Sdate$gender)
levels(Sdate$gender)<-c("female","Male")
ggplot(Sdate[which(!is.na(Sdate$age) & !is.na(Sdate$age_o)),], aes(x=as.factor(age), y=age_o, colour=gender)) +
    geom_boxplot(aes()) +
    geom_smooth(aes(group=as.factor(gender)), se=FALSE) + 
    labs(list(title="Age Range for Dating", x = "Age", y = "Age of Partner")) +
    ylim(18,40)
summary(Sdate$age)
    
```
# Explore race
```{r}
race.c <- c(
  "Black/African American",
  "European/Caucasian-American",
  "Latino/Hispanic American",
  "Asian/Pacific Islander/Asian-American",
  "Native American",
  "Other"
)

races <- Sdate[!is.na(Sdate$race),] %>%
  group_by(gender, race) %>%
  summarise(
    my.n = n()
  )

ggplot(races, aes(x = race, y = my.n, fill = factor(gender))) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_discrete(name = "Gender") +
  xlab("Race") + ylab("Count") + ggtitle("Race repartition") +
  scale_x_continuous(labels = race.c, breaks = 1:6) +
  coord_flip()
```
# Explore dating goal
```{r}
goal.c <- c(
  "Seemed like a fun night out",
  "To meet new people",
  "To get a date",
  "Looking for a serious relationship",
  "To say I did it",
  "Other"
)

goals <- Sdate[!is.na(Sdate$goal),] %>%
  group_by(gender, goal) %>%
  summarise(
    my.n = n()
  )

ggplot(goals, aes(x = goal, y = my.n, fill = factor(gender))) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_discrete(name = "Gender") +
  xlab("Goal") + ylab("Count") + ggtitle("Goal repartition") +
  scale_x_continuous(labels = goal.c, breaks = 1:6) +
  coord_flip()
```
# Explore normal date times 
```{r}
date.c <- c(
  "Several times a week",
  "Twice a week",
  "Once a week",
  "Twice a month",
  "Once a month",
  "Several times a year",
  "Almost never"
)

dates <- Sdate[!is.na(Sdate$date),] %>%
  group_by(gender, date) %>%
  summarise(
    my.n = n()
  )

ggplot(dates, aes(x = date, y = my.n, fill = factor(gender))) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_discrete(name = "Gender") +
  xlab("Date") + ylab("Count") + ggtitle("Date repartition") +
  scale_x_continuous(labels = date.c, breaks = 1:7) +
  coord_flip()
```

#number of missing values in each column
```{r}
sapply(Sdate11,function(x) sum(is.na(x)))
```
#number of unique values in each column
```{r}
sapply(Sdate11, function(x) length(unique(x)))
```
#Taking care of the missing values
A typical approach is to replace the missing values with the average, the median or the mode of the existing one. We use the average.
```{r}
Sdate11$attr[is.na(Sdate11$attr)] <- mean(Sdate11$attr,na.rm=T)
Sdate11$sinc[is.na(Sdate11$sinc)] <- mean(Sdate11$sinc,na.rm=T)
Sdate11$intel[is.na(Sdate11$intel)] <- mean(Sdate11$intel,na.rm=T)
Sdate11$fun[is.na(Sdate11$fun)] <- mean(Sdate11$fun,na.rm=T)
Sdate11$amb[is.na(Sdate11$amb)] <- mean(Sdate11$amb,na.rm=T)
Sdate11$shar[is.na(Sdate11$shar)] <- mean(Sdate11$shar,na.rm=T)

Sdate11$imprelig[is.na(Sdate11$imprelig)] <- mean(Sdate11$imprelig,na.rm=T)
Sdate11$goal[is.na(Sdate11$goal)] <- mean(Sdate11$goal,na.rm=T)
Sdate11$date[is.na(Sdate11$date)] <- mean(Sdate11$date,na.rm=T)
Sdate11$go_out[is.na(Sdate11$go_out)] <- mean(Sdate11$go_out,na.rm=T)
Sdate11$exphappy[is.na(Sdate11$exphappy)] <- mean(Sdate11$exphappy,na.rm=T)
Sdate11$age[is.na(Sdate11$age)] <- mean(Sdate11$age,na.rm=T)



library(Amelia)
missmap(Sdate11,main = "Missing values vs observed")

```
#Predictors selection
```{r}

library(psych)
pairs.panels(Sdate[c("dec", "attr","sinc","intel","fun","amb","shar","gender","age","imprelig","goal","date","go_out","exphappy")])
```



#Design an logistic regression analysis
We use dec as the response variable. Predictors are attr(attractive), sinc(sincere), intel(Intelligent), fun (Fun),	amb(ambitious),	shar(shared interests/hobbies).

First, we split the dataset into training and testing dataset.
```{r}
set.seed(123)
indx=sample(1:nrow(Sdate11), as.integer(0.9*nrow(Sdate11)))

Sdate11_train = Sdate11[indx,]
Sdate11_test = Sdate11[-indx,]

Sdate11_train_labels = Sdate11[indx,9]
Sdate11_test_labels = Sdate11[-indx,9]   

Sdate11_lm<-glm(dec~attr+sinc+intel+fun+amb+shar+gender+samerace+imprelig+goal+age+date+go_out+exphappy,family=binomial,data=Sdate11_train)
summary(Sdate11_lm)
```
We use 90% dataset as the training data and the left 10% as the testing data.

#Check training data and testing data randomness
```{r}
Sdata11_train_gender<-as.factor(Sdate11_train$gender)
summary(Sdata11_train_gender)
Sdata11_test_gender<-as.factor(Sdate11_test$gender)
summary(Sdata11_test_gender)
Sdata11_train_dec<-as.factor(Sdate11_train$dec)
summary(Sdata11_train_dec)
Sdata11_test_dec<-as.factor(Sdate11_test$dec)
summary(Sdata11_test_dec)
```


#AIC method
#Specifying the null model
```{r}
null=glm(dec~1, family=binomial(link='logit'),data=Sdate11_train)
summary(null)
full =glm(dec~attr+sinc+intel+fun+amb+shar+gender+samerace+imprelig+goal+age+date+go_out+exphappy, family=binomial(link='logit'),data= Sdate11_train)
summary(full)
step(null, scope=list(lower=null, upper=full), direction="forward")
```

Fit the logistic regression full model, with all predictor variables
```{r}
Sdate11_lm<-glm(dec~attr+sinc+intel+fun+amb+shar+gender+imprelig+ date+go_out+exphappy, family=binomial(link='logit'),data=Sdate11_train)
summary(Sdate11_lm)
```

#Check accuracy of the model
```{r}
glm.pred <- predict(Sdate11_lm,newdata=Sdate11_test,type='response')
glm.pred <- ifelse(glm.pred > 0.5,1,0)

table(glm.pred,Sdate11_test$dec)
misClasificError <- mean(glm.pred != Sdate11_test$dec)
print(paste('Accuracy',1-misClasificError))

```

#plot ROC
```{r}
library(ROCR)
p <- predict(Sdate11_lm, newdata=Sdate11_test, type="response")
pr <- prediction(p, Sdate11_test$dec)
prf <- performance(pr, measure = "tpr", x.measure = "fpr")
plot(prf)

auc <- performance(pr, measure = "auc")
auc <- auc@y.values[[1]]
auc

```


#Improve the model performance
drop the insignificant predictor, alpha = 0.05
From the regression model outputs, we can see that intel does not have a significant effect on decisions.
```{r}
Sdate12_lm<-glm(dec~attr + sinc + fun + amb + shar + 
    gender + imprelig + date + go_out, family=binomial(link='logit'),data=Sdate11_train)
summary(Sdate12_lm)

glm.pred <- predict(Sdate12_lm,newdata=Sdate11_test,type='response')
glm.pred <- ifelse(glm.pred > 0.5,1,0)

table(glm.pred,Sdate11_test$dec)
misClasificError <- mean(glm.pred != Sdate11_test$dec)
print(paste('Accuracy',1-misClasificError))
```

#plot ROC for the new model
```{r}
library(ROCR)
p <- predict(Sdate12_lm, newdata=Sdate11_test, type="response")
pr <- prediction(p, Sdate11_test$dec)
prf <- performance(pr, measure = "tpr", x.measure = "fpr")
plot(prf)

auc <- performance(pr, measure = "auc")
auc <- auc@y.values[[1]]
auc
```

We can see that the accuracy almost stay the same. So we keep the sincerity predictor in the model.????


#Correlation between preditors#########
```{r}
library(psych)
pairs.panels(Sdate11[c("dec", "attr","sinc","intel","fun","amb","shar","gender","age","imprelig","goal","date","go_out","exphappy")])

```

##Gender Differences
#Split the data into male and female
```{r}
Sdate3<-subset(Sdate,gender=="female")
Sdate4<-subset(Sdate,gender=="Male")
```

```{r}
which(colnames(Sdate)=="gender")
```
#Select the dataset we are interested in
```{r}
SdateF<-Sdate3[,c(3,15,34,42,46:48,68,98:104)]
SdateM<-Sdate4[,c(3,15,34,42,46:48,68,98:104)]
```

#Check the missing values
```{r}
missmap(SdateF)
missmap(SdateM)
```

#Taking care of the missing values
```{r}
SdateF$attr[is.na(SdateF$attr)] <- mean(SdateF$attr,na.rm=T)
SdateF$sinc[is.na(SdateF$sinc)] <- mean(SdateF$sinc,na.rm=T)
SdateF$intel[is.na(SdateF$intel)] <- mean(SdateF$intel,na.rm=T)
SdateF$fun[is.na(SdateF$fun)] <- mean(SdateF$fun,na.rm=T)
SdateF$amb[is.na(SdateF$amb)] <- mean(SdateF$amb,na.rm=T)
SdateF$shar[is.na(SdateF$shar)] <- mean(SdateF$shar,na.rm=T)

SdateF$imprelig[is.na(SdateF$imprelig)] <- mean(SdateF$imprelig,na.rm=T)
SdateF$goal[is.na(SdateF$goal)] <- mean(SdateF$goal,na.rm=T)
SdateF$date[is.na(SdateF$date)] <- mean(SdateF$date,na.rm=T)
SdateF$go_out[is.na(SdateF$go_out)] <- mean(SdateF$go_out,na.rm=T)
SdateF$exphappy[is.na(SdateF$exphappy)] <- mean(SdateF$exphappy,na.rm=T)
SdateF$age[is.na(SdateF$age)] <- mean(SdateF$age,na.rm=T)

SdateM$attr[is.na(SdateM$attr)] <- mean(SdateM$attr,na.rm=T)
SdateM$sinc[is.na(SdateM$sinc)] <- mean(SdateM$sinc,na.rm=T)
SdateM$intel[is.na(SdateM$intel)] <- mean(SdateM$intel,na.rm=T)
SdateM$fun[is.na(SdateM$fun)] <- mean(SdateM$fun,na.rm=T)
SdateM$amb[is.na(SdateM$amb)] <- mean(SdateM$amb,na.rm=T)
SdateM$shar[is.na(SdateM$shar)] <- mean(SdateM$shar,na.rm=T)


SdateM$imprelig[is.na(SdateM$imprelig)] <- mean(SdateM$imprelig,na.rm=T)
SdateM$goal[is.na(SdateM$goal)] <- mean(SdateM$goal,na.rm=T)
SdateM$date[is.na(SdateM$date)] <- mean(SdateM$date,na.rm=T)
SdateM$go_out[is.na(SdateM$go_out)] <- mean(SdateM$go_out,na.rm=T)
SdateM$exphappy[is.na(SdateM$exphappy)] <- mean(SdateM$exphappy,na.rm=T)
SdateM$age[is.na(SdateM$age)] <- mean(SdateM$age,na.rm=T)


library(Amelia)
missmap(SdateF,main = "Missing values vs observed")
missmap(SdateM,main = "Missing values vs observed")
```
#Fit the logiestic regression value
##Female
```{r}
set.seed(123)
indxF=sample(1:nrow(SdateF), as.integer(0.9*nrow(SdateF)))

SdateF_train = SdateF[indxF,]
SdateF_test = SdateF[-indxF,]

SdateF_train_labels = SdateF[indxF,2]
SdateF_test_labels = SdateF[-indxF,2] 

SdataF_train_dec<-as.factor(SdateF_train$dec)
summary(SdataF_train_dec)

SdateF_lm<-glm(dec~attr + sinc + intel + fun + amb + shar + imprelig + date + go_out + exphappy, family=binomial(link='logit'),data=SdateF_train)
summary(SdateF_lm)

glm.pred <- predict(SdateF_lm,newdata=SdateF_test,type='response')
glm.pred <- ifelse(glm.pred > 0.5,1,0)

table(glm.pred,SdateF_test$dec)
misClasificError <- mean(glm.pred != SdateF_test$dec)
print(paste('Accuracy',1-misClasificError))

library(ROCR)
pF <- predict(SdateF_lm, newdata=SdateF_test, type="response")
prF <- prediction(pF, SdateF_test$dec)
prfF <- performance(prF, measure = "tpr", x.measure = "fpr")
plot(prfF)

auc <- performance(prF, measure = "auc")
auc <- auc@y.values[[1]]
auc


```
#Improve the model 
drop insignificant predictor-intel
```{r}
SdateF2_lm<-glm(dec~attr + intel + fun + amb + shar + exphappy, family=binomial(link='logit'),data=SdateF_train)
summary(SdateF2_lm)

glm.pred <- predict(SdateF2_lm,newdata=SdateF_test,type='response')
glm.pred <- ifelse(glm.pred > 0.5,1,0)

table(glm.pred,SdateF_test$dec)
misClasificError <- mean(glm.pred != SdateF_test$dec)
print(paste('Accuracy',1-misClasificError))

library(ROCR)
pF2<- predict(SdateF2_lm, newdata=SdateF_test, type="response")
prF2 <- prediction(pF2, SdateF_test$dec)
prfF2 <- performance(prF2, measure = "tpr", x.measure = "fpr")
plot(prfF2)

auc <- performance(prF2, measure = "auc")
auc <- auc@y.values[[1]]
auc
```

##Male
```{r}
set.seed(123)
indxM=sample(1:nrow(SdateM), as.integer(0.9*nrow(SdateM)))

SdateM_train = SdateM[indxM,]
SdateM_test = SdateM[-indxM,]

SdateM_train_labels = SdateM[indxM,2]
SdateM_test_labels = SdateM[-indxM,2]   

SdataM_test_dec<-as.factor(SdateM_test$dec)
summary(SdataM_test_dec)

SdateM_lm<-glm(dec~attr + sinc + intel + fun + amb + shar + imprelig + date + go_out + exphappy, family=binomial(link='logit'),data=SdateM_train)
summary(SdateM_lm)

glm.pred <- predict(SdateM_lm,newdata=SdateM_test,type='response')
glm.pred <- ifelse(glm.pred > 0.5,1,0)

table(glm.pred,SdateM_test$dec)
misClasificError <- mean(glm.pred != SdateM_test$dec)
print(paste('Accuracy',1-misClasificError))

library(ROCR)
p <- predict(SdateM_lm, newdata=SdateM_test, type="response")
pr <- prediction(p, SdateM_test$dec)
prf <- performance(pr, measure = "tpr", x.measure = "fpr")
plot(prf)

auc <- performance(pr, measure = "auc")
auc <- auc@y.values[[1]]
auc
```
#Improve the model 
drop insignificant predictor-intel
```{r}
SdateM2_lm<-glm(dec~attr + sinc + fun + amb + shar + imprelig + date + exphappy, family=binomial(link='logit'),data=SdateM_train)
summary(SdateM2_lm)

glm.pred <- predict(SdateM2_lm,newdata=SdateM_test,type='response')
glm.pred <- ifelse(glm.pred > 0.5,1,0)

table(glm.pred,SdateM_test$dec)
misClasificError <- mean(glm.pred != SdateM_test$dec)
print(paste('Accuracy',1-misClasificError))

library(ROCR)
pF2<- predict(SdateM2_lm, newdata=SdateM_test, type="response")
prF2 <- prediction(pF2, SdateM_test$dec)
prfF2 <- performance(prF2, measure = "tpr", x.measure = "fpr")
plot(prfF2)

auc <- performance(prF2, measure = "auc")
auc <- auc@y.values[[1]]
auc
```


